### 2023-11-30 state/crossbar.tex
Added almost all symbols

It uses physical properties of electrical systems to perform analog computing. In what follows I will use the circuit node in \cref{fig:crossNode} to explain the theory behind the memristor crossbar array.
\begin{figure}[H]
  \centering
  \includesvg[height=3.5cm]{crossbar/node}
  %\def\svgheigth{3.5cm}
  %\input{crossbar/node.pdf_tex}
  \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}
  \label{fig:crossNode}
\end{figure}

First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $R_i$, is $V_i$. As such and using Ohm's law, we know that the current flowing into the memristor ($I_{i}$) is bound by the following equation :

\begin{equation}
  V_i = R\cdot I_{i} \Rightarrow I_{i} = V_i\cdot (\frac{1}{R})= V_i\cdot G_i
\end{equation}
With $G_i$ being the conductance of memristor, defined as $ G_i=\frac{1}{R_i}$.

This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
\begin{equation}
  I_{j,i} = I_{j,i-1}+I_{i} = I_{j,i-1} + V_i\cdot G_i
\end{equation}
By applying this process to the whole system we find that the current at the bottom of one column is :
\begin{equation}
  I_1= G_1\cdot V_1 +  G_2\cdot V_2 +  G_3\cdot V_3
\end{equation}
With $G_1$, $G_2$ and $G_3$ being the conductance of the 3 memristors in the first column.

>>>>

It uses physical properties of electrical systems to perform analog computing. In what follows I will use the circuit node in \cref{fig:crossNode} to explain the theory behind the memristor crossbar array.
\begin{figure}[H]
  \centering
  \includesvg[height=3.5cm]{crossbar/node}
  %\def\svgheigth{3.5cm}
  %\input{crossbar/node.pdf_tex}
  \caption{Memristor crossbar node of the $k^{th}$ line and $j^{th}$ column}
  \label{fig:crossNode}
\end{figure}

First of all, a voltage is applied on the $k^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $\symR_k$, is $\symv_k$. As such and using Ohm's law, we know that the current flowing into the memristor ($\symi_{k}$) is bound by the following equation :

\begin{equation}
  \symv_k = \symR_k\cdot \symi_{k} \Rightarrow \symi_{k} = \symv_k\cdot (\frac{1}{\symR_k})= \symv_k\cdot \symG_k
\end{equation}
With $\symG_k$ being the conductance of memristor, defined as $\symG_k=\frac{1}{\symR_k}$.

This line then joins the column where a current of $\symi_{j,k-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
\begin{equation}
  \symi_{j,k} = \symi_{j,k-1}+\symi_{k} = \symi_{j,k-1} + \symv_k\cdot \symG_k
\end{equation}
By applying this process to the whole system we find that the current at the bottom of one column is :
\begin{equation}
  \symi_1= \symG_1\cdot \symv_1 +  \symG_2\cdot \symv_2 +  \symG_3\cdot \symv_3
\end{equation}
With $\symG_1$, $\symG_2$ and $\symG_3$ being the conductance of the 3 memristors in the first column.

### 2023-11-30 state/gru.tex
Correctiion diogo state
Changed a bunch of files
Added almost all symbols

\section{\acs{GRU}}\label{sec:gru}
The \acf{GRU} is another type of \ac{RNN}. It is also known to reduce the effect of the vanishing gradient problem. It was first introduced to improve translation techniques \cite{gru}.

@@

\begin{equation}\label{eq:updateG}
  \overrightarrow{z_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_z + \overrightarrow{b_z})
\end{equation}
\begin{equation}\label{eq:resetG}
  \overrightarrow{r_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_r + \overrightarrow{b_r})
\end{equation}
\begin{equation}\label{eq:candActivG}
  \overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+(\overrightarrow{r_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{hh} + \overrightarrow{b_h})
\end{equation}
\begin{equation}\label{eq:gruHidG}
  \overrightarrow{h_t}=(\overrightarrow{1}-\overrightarrow{z_t})\odot \overrightarrow{h_{t-1}} + \overrightarrow{z_t}\odot \overrightarrow{\hat{h_t}}
\end{equation}

Where ($w_z$,$\overrightarrow{b_z}$), ($w_r$,$\overrightarrow{b_r}$),($w_{hx}$,$w_{hh}$,$\overrightarrow{b_h}$) are the weights matrixes and bias vectors for the update, reset and candidate activation gates respectively.

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/encoderCell.svg}
  \label{fig:encoderGruCell}
  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}}
\end{figure}

@@

\begin{equation}\label{eq:gruHidG1}
  \overrightarrow{h_t}=\overrightarrow{z_t}\odot \overrightarrow{h_{t-1}} + (\overrightarrow{1}-\overrightarrow{z_t})\odot \overrightarrow{\hat{h_t}}
\end{equation}

@@

\begin{equation}\label{eq:candActivG1}
  \overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+ \overrightarrow{b_{hx}}+\overrightarrow{r_t}\odot[\overrightarrow{h_{t-1}} \cdot w_{hh} + \overrightarrow{b_{hh}}])
\end{equation}

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/decoderCell.svg}
  \label{fig:decoderGruCell}
  \caption{Decoder \acs{GRU} cell, legend in \cref{leg:cells}}
\end{figure}

@@

\begin{equation}\label{eq:lstmGru0}
  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_f + \overrightarrow{b_f})
\end{equation}
\begin{equation}\label{eq:lstmGru1}
  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_o + \overrightarrow{b_o})
\end{equation}
\begin{equation}\label{eq:lstmGru2}
  \overrightarrow{i_t}=\overrightarrow{1}-\overrightarrow{f_t}
\end{equation}
\begin{equation}\label{eq:lstmGru3}
  \overrightarrow{\tilde{c_t}}=tanh(\overrightarrow{x_t}\cdot w_{cx}+(\overrightarrow{o_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{ch} + \overrightarrow{b_c})
\end{equation}
\begin{equation}\label{eq:lstmGru4}
  \overrightarrow{h_t}=(\overrightarrow{i_t})\odot \overrightarrow{h_{t-1}} + \overrightarrow{f_t}\odot \overrightarrow{\tilde{h_t}}
\end{equation}

>>>>

\section{\acl{GRU}}\label{sec:gru}
The \acf{GRU} is another type of \ac{RNN}. It is also known to reduce the effect of the vanishing gradient problem. It was first introduced to improve translation techniques \cite{gru}.

@@

\begin{equation}\label{eq:updateG}
  \symvz_t=\sigma ([\symvx_t,\symvh_{t-1}] \cdot \symmw_z + \symvb_z)
\end{equation}
\begin{equation}\label{eq:resetG}
  \symvr_t=\sigma ([\symvx_t,\symvh_{t-1}] \cdot \symmw_r + \symvb_r)
\end{equation}
\begin{equation}\label{eq:candActivG}
  \symvhh_t=tanh(\symvx_t\cdot \symmw_{hx}+(\symvr_t\odot\symvh_{t-1}) \cdot \symvw_{hh} + \symvb_h)
\end{equation}
\begin{equation}\label{eq:gruHidG}
  \symvh_t=(\vunit-\symvz_t)\odot \symvh_{t-1} + \symvz_t\odot \symvhh_t
\end{equation}

Where ($\symmw_z$,$\symvb_z$), ($\symmw_r$,$\symvb_r$),($\symmw_{hx}$,$\symmw_{hh}$,$\symvb_h$) are the weights matrixes and bias vectors for the update, reset and candidate activation gates respectively.

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/encoderCell.svg}
  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}}
  \label{fig:encoderGruCell}
\end{figure}

@@

\begin{equation}\label{eq:gruHidG1}
  \symvh_t=\symvz_t\odot \symvh_{t-1} + (\vunit-\symvz_t)\odot \symvh_t
\end{equation}

@@

\begin{equation}\label{eq:candActivG1}
  \symvhh_t=tanh(\symvx_t\cdot \symmw_{hx}+ \symvb_{hx}+\symvr_t\odot[\symvh_{t-1} \cdot \symmw_{hh} + \symvb_{hh}])
\end{equation}

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/decoderCell.svg}
  \caption{Decoder \acs{GRU} cell, legend in \cref{leg:cells}}
  \label{fig:decoderGruCell}
\end{figure}

@@

\begin{equation}\label{eq:lstmGru0}
  \symvf_t=\sigma ([\symvx_t,\symvh_{t-1}] \cdot \symmw_f + \symvb_f)
\end{equation}
\begin{equation}\label{eq:lstmGru1}
  \symvo_t=\sigma ([\symvx_t,\symvh_{t-1}] \cdot \symmw_o + \symvb_o)
\end{equation}
\begin{equation}\label{eq:lstmGru2}
  \symvi_t=\vunit-\symvf_t
\end{equation}
\begin{equation}\label{eq:lstmGru3}
  \symvct_t=tanh(\symvx_t\cdot \symmw_{cx}+(\symvo_t\odot\symvh_{t-1}) \cdot \symmw_{ch} + \symvb_c)
\end{equation}
\begin{equation}\label{eq:lstmGru4}
  \symvh_t=(\symvi_t)\odot \symvh_{t-1} + \symvf_t\odot \symvct_t
\end{equation}

### 2023-11-30 state/lstm.tex
Added almost all symbols

\begin{equation}\label{eq:inputG}
  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i})
\end{equation}
\begin{equation}\label{eq:forgetG}
  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f})
\end{equation}
\begin{equation}\label{eq:outputG}
  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o})
\end{equation}

Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$) and ($w_o$,$\overrightarrow{b_o}$) are the pair of weights matrixes and bias vectors for the input, forget and output gates respectively. $\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.

@@

\begin{equation}\label{eq:candCell}
  \overrightarrow{\tilde{c}_t}=tanh([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_c+ \overrightarrow{b_c})
\end{equation}
\begin{equation}\label{eq:cellS}
  \overrightarrow{c_t}=\overrightarrow{f_t}\odot \overrightarrow{c_{t-1}} + \overrightarrow{i_t} \odot \overrightarrow{\tilde{c}_t}
\end{equation}

Where $w_c$ and $\overrightarrow{b_c}$ are the weights matrix and bias vector for the candidate cell state.

The final step of the \ac{LSTM} is to compute the hidden state (\cref{eq:hiddenS}).
\begin{equation}\label{eq:hiddenS}
  \overrightarrow{h_t}=\overrightarrow{o_t}\odot tanh(\overrightarrow{c_t})
\end{equation}

We set $x_1$ as the first input and define $\overrightarrow{h_0}$ as a zero only vector.

@@

\begin{equation}\label{eq:inputGVanilla}
  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
\end{equation}
\begin{equation}\label{eq:forgetGVanilla}
  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f}+\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
\end{equation}
\begin{equation}\label{eq:ouputGVanilla}
  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o}+\overrightarrow{c_{t}}\odot \overrightarrow{p_o})
\end{equation}

With $\overrightarrow{p_f}$, $\overrightarrow{p_i}$ and $\overrightarrow{p_o}$ being the peephole weights vectors. Their size is the same as the size of the hidden state vector ($\overrightarrow{h_t}$). $\odot$ is the pointwise multiplication operator.

The equations that have not been rewritten simply stay the same.

@@

\begin{equation}\label{eq:inputGFGR}
  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_i  + \overrightarrow{i_{t-1}}\cdot R_{ii} + \overrightarrow{f_{t-1}}\cdot R_{if} + \overrightarrow{o_{t-1}}\cdot R_{io} + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
\end{equation}
\begin{equation}\label{eq:forgetGFGR}
  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_f  + \overrightarrow{i_{t-1}}\cdot R_{fi} + \overrightarrow{f_{t-1}}\cdot R_{ff} + \overrightarrow{o_{t-1}}\cdot R_{fo} + \overrightarrow{b_f} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
\end{equation}
\begin{equation}\label{eq:ouputGFGR}
  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{i_{t-1}}\cdot R_{oi} + \overrightarrow{f_{t-1}}\cdot R_{of} + \overrightarrow{o_{t-1}}\cdot R_{oo} + \overrightarrow{b_o} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_o})
\end{equation}

With $R_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrixes.

@@

\begin{itemize}
  \item Removing the activation function from the Vanilla \ac{LSTM} for :
    \begin{itemize}
      \item the candidate cell and thus becomes $ \overrightarrow{\tilde{c}_t}=(w_c\cdot[\overrightarrow{x_{t-1}},\overrightarrow{h_{t-1}}] + \overrightarrow{b_c}) $, this is called the \ac{NIAF}
      \item the output of the previous cell state and thus becomes $ \overrightarrow{h_t}=\overrightarrow{o_t}\odot \overrightarrow{c_t} $, and is called \ac{NOAF}
    \end{itemize}
  \item Using units instead of the respective gates :
    \begin{itemize}
      \item \ac{NIG} : $\overrightarrow{i_t}=1$
      \item \ac{NOG} : $\overrightarrow{o_t}=1$
      \item \ac{NFG} : $\overrightarrow{f_t}=1$
    \end{itemize}
\end{itemize}

>>>>

\begin{equation}\label{eq:inputG}
  \symvi_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_i + \symvb_i)
\end{equation}
\begin{equation}\label{eq:forgetG}
  \symvf_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_f + \symvb_f)
\end{equation}
\begin{equation}\label{eq:outputG}
  \symvo_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_o + \symvb_o)
\end{equation}

Where ($\symmw_i$,$\symvb_i$), ($\symmw_f$,$\symvb_f$) and ($\symmw_o$,$\symvb_o$) are the pair of weights matrixes and bias vectors for the input, forget and output gates respectively. $\symvx_t$ is the input vector and $\symvh_t$ is the hidden state vector.

@@

\begin{equation}\label{eq:candCell}
  \symvct_t=tanh([\symvx_{t},\symvh_{t-1}]\cdot \symmw_c+ \symvb_c)
\end{equation}
\begin{equation}\label{eq:cellS}
  \symvc_t=\symvf_t\odot \symvc_{t-1} + \symvi_t \odot \symvct_t
\end{equation}

Where $\symmw_c$ and $\symvb_c$ are the weights matrix and bias vector for the candidate cell state.

The final step of the \ac{LSTM} is to compute the hidden state (\cref{eq:hiddenS}).
\begin{equation}\label{eq:hiddenS}
  \symvh_t=\symvo_t\odot tanh(\symvc_t)
\end{equation}

We set $x_1$ as the first input and define $\symvh_0$ as a zero only vector.

@@

\begin{equation}\label{eq:inputGVanilla}
  \symvi_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_i + \symvb_i +\symvc_{t-1}\odot \symvp_f)
\end{equation}
\begin{equation}\label{eq:forgetGVanilla}
  \symvf_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_f + \symvb_f+\symvc_{t-1}\odot \symvp_i)
\end{equation}
\begin{equation}\label{eq:ouputGVanilla}
  \symvo_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_o + \symvb_o+\symvc_{t}\odot \symvp_o)
\end{equation}

With $\symvp_f$, $\symvp_i$ and $\symvp_o$ being the peephole weights vectors. Their size is the same as the size of the hidden state vector ($\symvh_t$).
The equations that have not been rewritten simply stay the same.

@@

\begin{equation}\label{eq:inputGFGR}
  \symvi_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_i  + \symvi_{t-1}\cdot \symmw_{ii} + \symvf_{t-1}\cdot \symmw_{if} + \symvo_{t-1}\cdot \symmw_{io} + \symvb_i +\symvc_{t-1}\odot \symvp_i)
\end{equation}
\begin{equation}\label{eq:forgetGFGR}
  \symvf_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_f  + \symvi_{t-1}\cdot \symmw_{fi} + \symvf_{t-1}\cdot \symmw_{ff} + \symvo_{t-1}\cdot \symmw_{fo} + \symvb_f +\symvc_{t-1}\odot \symvp_f)
\end{equation}
\begin{equation}\label{eq:ouputGFGR}
  \symvo_t=\sigma ([i\symvx_{t},\symvh_{t-1}]\cdot \symmw_o + \symvi_{t-1}\cdot \symmw_{oi} + \symvf_{t-1}\cdot \symmw_{of} + \symvo_{t-1}\cdot \symmw_{oo} + \symvb_o +\symvc_{t-1}\odot \symvp_o)
\end{equation}

With $\symmw_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrixes.

@@

\begin{itemize}
  \item Removing the activation function from the Vanilla \ac{LSTM} for :
    \begin{itemize}
      \item the candidate cell and thus becomes $ \symvct_t=([\symvx_{t-1},\symvh_{t-1}]\cdot\symmw_c + \symvb_c) $, this is called the \ac{NIAF}
      \item the output of the previous cell state and thus becomes $ \symvh_t=\symvo_t\odot \symvc_t $, and is called \ac{NOAF}
    \end{itemize}
  \item Using units instead of the respective gates :
    \begin{itemize}
      \item \ac{NIG} : $\symvi_t=1$
      \item \ac{NOG} : $\symvo_t=1$
      \item \ac{NFG} : $\symvf_t=1$
    \end{itemize}
\end{itemize}

### 2023-11-30 state/memristors.tex
Correctiion diogo state
Added almost all symbols

Memristors are the lesser known fourth fundamental passive component of electronics, among resistors, capacitors and inductor.
It was first theorized in 1971 by L. Chua from UC Berkeley, in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
The theory behind this component was extracted from a missing component to link the four fundamental circuit variables, voltage ($v$), charge ($q$), current ($i$) and flux ($\phi$). \Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :

\begin{equation}
  d\phi = v\cdot dt
\end{equation}

\begin{equation}
  dq = i\cdot dt
\end{equation}

Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux ($\phi$) and charge ($q$).  The flux in this case is not a magnetic flux and is defined as such : $ d\phi=v\cdot dt \Rightarrow \phi =  \int v \,dt  $.

@@

\begin{figure}[H]
  \centering
  \includesvg[width=0.55\textwidth]{memristor/memristor}
  \caption{Fundamental passive components}
  \label{fig:fundComp}
\end{figure}

\subsection{Equations}
A memristor links the flux ($\phi$) and charge ($q$) by the memristance.
This memristance is defined with the following equation :
\begin{equation}
  M(q)=\frac{d\phi}{dq}
\end{equation}
It can be compared with the other fundamental components like resistor ($R(i)=\frac{dv}{di}$), capacitor ($\frac{1}{C(q)}=\frac{dv}{dq}$) and inductor ($L(i)=\frac{d\phi}{di}$).
We can then extract a more useful equation in an actual circuit :
\begin{equation}
  v(t)=M(q(t))\cdot i(t)
\end{equation}
Similarly, a memductance can be defined as such :
\begin{equation}
  W(\phi)=\frac{dq}{d\phi}
\end{equation}
A memristive device is slightly differently defined, it still uses a memristance, but here the memristance also depends on an internal state called $x$. This gives us this equation :
\begin{equation}\label{eq:memristiveDev}
  v(t)=M(x,i)\cdot i(t)
\end{equation}
The internal state ($x$) is not linked to flux or charge in the case of a memristive device.\\
Once again, we can also define the memristive device using a memductance :
\begin{equation}
  i(t)=W(x,v)\cdot v(t)
\end{equation}
In all of the previous equations, $v$ is the voltage in Volt ($V$), $i$ is the current in Ampere ($A$), $\phi$ is the flux in Weber ($Wb$), $q$ is the charge in Coulomb ($C$), $M$ is the memristance in Ohm ($\Omega$) and $W$ is the memductance in Siemens ($S$ or $\Omega^{-1}$).

>>>>

Memristors are the lesser known fourth fundamental passive component of electronics, among resistors, capacitors and inductor.
It was first theorized in 1971 by L. Chua from UC Berkeley, in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
The theory behind this component was extracted from a missing component to link the four fundamental circuit variables, voltage (\symv), charge (\symq), current (\symq) and flux (\symphi). \Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :

\begin{equation}
  d\symphi = \symv\cdot d\symt
\end{equation}

\begin{equation}
  d\symq = \symi\cdot d\symt
\end{equation}

Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux (\symphi) and charge (\symq).  The flux in this case is not a magnetic flux and is defined as such : $d\symphi=\symv\cdot d\symt \Rightarrow \symphi =  \int \symv \,d\symt$.

@@

\begin{figure}[H]
  \centering
  \includesvg[width=0.55\textwidth]{memristor/memristor}
  \caption{Fundamental passive components, adapted from \cite{memWiki}}
  \label{fig:fundComp}
\end{figure}

\subsection{Equations}
A memristor links the flux (\symphi) and charge (\symq) by the memristance.
This memristance is defined with the following equation :
\begin{equation}
  \symM(\symq)=\frac{d\symphi}{d\symq}
\end{equation}
It can be compared with the other fundamental components like resistor ($\symR(\symi)=\frac{d\symv}{d\symi}$), capacitor ($\frac{1}{\symC(\symq)}=\frac{d\symv}{d\symq}$) and inductor ($\symL(\symi)=\frac{d\symphi}{d\symi}$).
We can then extract a more useful equation in an actual circuit :
\begin{equation}
  \symv(\symt)=\symM(\symq(\symt))\cdot \symi(\symt)
\end{equation}
Similarly, a memductance can be defined as such :
\begin{equation}
  \symW(\symphi)=\frac{d\symq}{d\symphi}
\end{equation}
A memristive device is slightly differently defined, it still uses a memristance, but here the memristance also depends on an internal state called $x$. This gives us this equation :
\begin{equation}\label{eq:memristiveDev}
  \symv(\symt)=\symM(x,\symi)\cdot \symi(\symt)
\end{equation}
The internal state ($x$) is not linked to flux or charge in the case of a memristive device.\\
Once again, we can also define the memristive device using a memductance :
\begin{equation}
  \symi(\symt)=\symW(x,\symv)\cdot \symv(\symt)
\end{equation}
In all of the previous equations, \symv is the voltage in Volt ($V$), \symi is the current in Ampere ($A$), \symphi is the flux in Weber ($Wb$), \symq is the charge in Coulomb ($C$), \symM is the memristance in Ohm ($\Omega$) and \symW is the memductance in Siemens ($S$ or $\Omega^{-1}$).

### 2023-11-30 state/model.tex
Removing space
Changed a bunch of files
Added almost all symbols

\begin{equation}
  \frac{dx}{dt}=f(x,v)
\end{equation}

Where $v$ is the voltage, $t$ is time and $f$ is a function.

@@

\begin{equation}\label{eq:funcDesc}
  f(x,v)=g(x)\cdot s(v)
\end{equation}

@@

\begin{equation}\label{eq:memModel0}
  i(R,v)=
  \begin{cases}
    \frac{a_p}{R}\cdot sinh(b_p\cdot v) & v\ge 0\\
    \frac{a_n}{R}\cdot sinh(b_n\cdot v) & v<0\\
  \end{cases}
\end{equation}

\begin{equation}\label{eq:memModel1}
  \frac{dR}{dt}=f(R,v)=s(v)\cdot g(R,v)
\end{equation}

@@

\begin{equation}\label{eq:memModel2}
  s(v)=
  \begin{cases}
    A_p\cdot (e^{t_p\cdot |v|}-1)& v>0\\
    A_n\cdot (e^{t_n\cdot |v|}-1)& v<0\\
    0 &  v=0\\
  \end{cases}
\end{equation}

@@

\begin{equation}\label{eq:memModel3}
  g(R,v)=
  \begin{cases}
    (R-r_p(v))^2& v>0\\
    (R-r_n(v))^2& v<0\\
    0 &  v=0\\
  \end{cases}
\end{equation}

@@

\begin{equation}\label{eq:memModel4}
  \begin{cases}
    r_p(v)= r_{p,0} + r_{p,1}\cdot v& v>0\\
    r_n(v)= r_{r,0} + r_{n,1}\cdot v& v\le 0\\
  \end{cases}
\end{equation}

@@

The changes of resistive states from the initial resistance ($R_0$) are computed by \cref{eq:memModel5} extracted from \cite{memCadenceModel}. It uses the bias voltage ($V_b$) of the pulse applied to the memristor. The pulse has a duration of $t$.

@@

The final Verilog-A code is available at \cref{apsec:memModel}.

@@

The model implements, as with physical memristors, a threshold voltage value from which its resistive states starts changing. This threshold value is $V_{read}$. For this specific model the threshold voltage is set to $V_{read}=500mV$.

Every pulse is composed of 2 parts and lasts for a total of $1 ms$ :
\begin{itemize}
  \item The reading of the resistive state, by applying a triangular wave that is bound by the threshold value $V_{read}$. The triangular wave starts from $0V$, goes to $V_{read}$ then to $-V_{read}$ before going back to $0V$.
  \item The writing to the memristor. During this part the resistive state will change. This is done by applying a square wave of variable width ($t_{\Delta}$) with a set bias voltage ($V_{bias}$).
\end{itemize}

@@

The resistive state of the memristor changes depending on the pulse width ($t_{\Delta}$) and the bias voltage ($V_{bias}$) used for each pulse.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memristor/memPulseChange.svg}
  \caption{Memristor's resistive state under different pulses width (Fixed bias voltage $V_{bias}=1.8V$)}
  \label{graph:pulseChange}
\end{figure}

The effect of the pulse width variation is shown in \cref{graph:pulseChange}. Indeed, the larger the pulse widths, the faster the internal resistance reaches the bounding resistance (\cref{eq:memModel4}). The bounding resistance being the same as all curves use the same bias voltage ($V_{bias}=1.8V$), the curves will all eventually reach the max resistance.

@@

The voltage dependence of the memristor's model is described in \cref{graph:voltChange}. The influence of the bias voltage is, as \cref{eq:memModel3,eq:memModel4} show, the way to change the resistance range of the memristor, until the physical limit is reached of course. Applying higher voltage also allows to reach a set resistance faster as the slopes are bigger the higher the bias voltage gets.

The resistive state can also be lowered by applying negative voltages.%TODO

>>>>

\begin{equation}
  \frac{dx}{d\symt}=f(x,\symv)
\end{equation}

Where \symv is the voltage, \symt is time.

@@

\begin{equation}\label{eq:funcDesc}
  f(x,\symv)=g(x)\cdot s(\symv)
\end{equation}

@@

\begin{equation}\label{eq:memModel0}
  \symi(\symR,\symv)=
  \begin{cases}
    \frac{a_p}{\symR}\cdot sinh(b_p\cdot \symv) & \symv\ge 0\\
    \frac{a_n}{\symR}\cdot sinh(b_n\cdot \symv) & \symv<0\\
  \end{cases}
\end{equation}

\begin{equation}\label{eq:memModel1}
  \frac{d\symR}{d\symt}=f(\symR,\symv)=s(\symv)\cdot g(\symR,\symv)
\end{equation}

@@

\begin{equation}\label{eq:memModel2}
  s(\symv)=
  \begin{cases}
    A_p\cdot (e^{t_p\cdot |v|}-1)& v>0\\
    A_n\cdot (e^{t_n\cdot |v|}-1)& v<0\\
    0 &  v=0\\
  \end{cases}
\end{equation}

@@

\begin{equation}\label{eq:memModel3}
  g(\symR,\symv)=
  \begin{cases}
    (\symR-r_p(\symv))^2& \symv>0\\
    (\symR-r_n(\symv))^2& \symv<0\\
    0 &  \symv=0\\
  \end{cases}
\end{equation}

@@

\begin{equation}\label{eq:memModel4}
  \begin{cases}
    r_p(\symv)= r_{p,0} + r_{p,1}\cdot \symv& \symv>0\\
    r_n(\symv)= r_{r,0} + r_{n,1}\cdot \symv& \symv\le 0\\
  \end{cases}
\end{equation}

@@

The changes of resistive states from the initial resistance ($\symR_0$) are computed by \cref{eq:memModel5} extracted from \cite{memCadenceModel}. It uses the bias voltage ($\symv_b$) of the pulse applied to the memristor. The pulse has a duration of $t$.

@@

The final Verilog-A code is available in \cite{memCadenceModel}.

@@

The model implements, as with physical memristors, a threshold voltage value from which its resistive states starts changing. This threshold value is $\symv_{read}$. For this specific model the threshold voltage is set to $\symv_{read}=500mV$.

Every pulse is composed of 2 parts and lasts for a total of $1 ms$ :
\begin{itemize}
  \item The reading of the resistive state, by applying a triangular wave that is bound by the threshold value $\symv_{read}$. The triangular wave starts from $0V$, goes to $\symv_{read}$ then to $-\symv_{read}$ before going back to $0V$.
  \item The writing to the memristor. During this part the resistive state will change. This is done by applying a square wave of variable width ($\symt_{\Delta}$) with a set bias voltage ($\symv_{bias}$).
\end{itemize}

@@

The resistive state of the memristor changes depending on the pulse width ($\symt_{\Delta}$) and the bias voltage ($\symv_{bias}$) used for each pulse.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memristor/memPulseChange.svg}
  \caption{Memristor's resistive state under different pulses width (Fixed bias voltage $\symv_{bias}=\symv_{dd}=1.8V$)}
  \label{graph:pulseChange}
\end{figure}

The effect of the pulse width variation is shown in \cref{graph:pulseChange}. Indeed, the larger the pulse widths, the faster the internal resistance reaches the bounding resistance (\cref{eq:memModel4}). The bounding resistance being the same as all curves use the same bias voltage ($\symv_{bias}=\symv_{dd}=1.8V$), the curves will all eventually reach the max resistance.

@@

The voltage dependence of the memristor's model is described in \cref{graph:voltChange}. The influence of the bias voltage is, as \cref{eq:memModel3,eq:memModel4} show, the way to change the resistance range of the memristor, until the physical limit is reached of course. Applying higher voltage also allows to reach a set resistance faster as the slopes are bigger the higher the bias voltage gets. The resistive state can also be lowered by applying negative voltages.%TODO
